---
title: "Econ573_ProblemSet1_PartIII"
output:
  pdf_document: default
  html_notebook: default
---

1.  Exploring Data

    prices are elevated in all regions, and we see that brownies flavor is the most sold flavor.

```{r}
barplot(table(benjer$flavor_descr), border = NA, las = 2)
```

```{r}
hist(y)
hist(priceper1)
plot(y ~ benjer$region)
```

2.  

3.  Regression Model

    ```{r}
    summary(fit)
    ```

    The refine the model, we can remove the flavor_descr variable to simplify the model. We can also remove other non-significant variables like race, hispanic_origin, and dishawasher.

    ```{r}
    fit2 <- glm(y ~ . - flavor_descr, data = x)
    summary(fit2)
    fit3 <- glm(y ~ . - flavor_descr - race - dishwasher - hispanic_origin, data = x)
    summary(fit3)

    ```

    Now we compare the models using AIC/BIC to see which one best fits the data.

    ```{r}
    n <- nrow(x)
    BIC <- c(
      reg1 = extractAIC(fit, k = log(n))[2],
      reg2 = extractAIC(fit2, k = log(n))[2],
      reg3 = extractAIC(fit3, k = log(n))[2]
    )

    # Model probabilities
    eBIC <- exp(-0.5 * (BIC - min(BIC)))
    probs <- eBIC / sum(eBIC)
    round(probs, 5)

    ```

    The simpler model (3) is better.

<!-- -->

3.  d

    ```{r}
    summary(fit)
    ```

Significant variables include: (Intercept), flavor_descrCINNAMON BUNS, flavor_descrCOFFEE, flavor_descrCREME BRULEE, flavor_descrHEATH CRUNCH, flavor_descrMINT CHC CHUNK, flavor_descrONE CSK BROWNIE, flavor_descrPUMPKIN CSK, flavor_descrSMORES, flavor_descrTURTLE SOUP, flavor_descrVAN, size1_descr32.0 MLOZ, household_income, household_size, usecoupTRUE, couponper1, regionCentral, regionSouth, regionWest, marriedTRUE, raceblack, raceother, microwaveTRUE, sfhTRUE, internetTRUE, tvcableTRUE
